---
title: "Peggoty_Loggers"
author: "Eva Gerstle"
date: "2024-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Step 2

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(VulnToolkit)
```

Step 3 & 4 Listing the file names of each logger's data

```{r}
pwd <- "/Users/evagerstle/Downloads/Calibrated_Data_PEG"
PEG_files <- list.files(path = pwd, pattern = ".csv")
```

Step 5 & 6 Reading in the elevation datasheet

```{r}
PEGelev <- read_xlsx("/Users/evagerstle/Downloads/03Aug2023_PEG_ArrayDataSheet.xlsx")
PEGelev <- PEGelev %>% 
  select(Serial, Type, Elevation)
  
```

Step 7 Create empty data frame

```{r}
flood_table_PEG = tibble('Serial_Number'= 0, 'Proportion_Time_Innudated'= 0, 'Median_Time_Innudated' = 0, 'Median_Depth_Innudated' = 0, 'Logger_Elevation' = 0, 'Type' = "", 'percentile_80_depth' = 0, 'percentile_95_depth' = 0, 'wet_75_percentile_depth' = 0)
```


 For loop where the magic happens

```{r}
for (file1 in PEG_files) {
  serialnum = as.integer(substring(file1, 1, 8))
  file1.adj = paste(pwd,file1, sep = "/")

  openedfile <- read_csv(file1.adj, col_names = TRUE, skip=1)

  logger_PEGelev_df <- PEGelev %>%
    filter(Serial == serialnum) %>%
   select(Elevation, Type)
   logger_PEGelev = as.numeric(logger_PEGelev_df[1])
    logger_PEGtype = as.character(logger_PEGelev_df[2])

     
 
 colnames(openedfile)[6] = "sensors"
  colnames(openedfile)[2] = "Date"
  colnames(openedfile)[1] = "#"
   openedfile <- openedfile %>%
  drop_na(sensors)
   
   
    openedfile <- openedfile %>%
     select(Date, `#`, sensors) %>%
     mutate(fixed_sensor = if_else(sensors < 0, 0, sensors)) %>%
    mutate(ht_above_SL = fixed_sensor + logger_PEGelev ) %>%
   mutate(DT2 = strptime(Date,format="%m/%d/%y %I:%M:%S %p"))
   
  
 at_loggerelev_freq <- fld.dur(logger_PEGelev + .02, openedfile$ht_above_SL)
 timeinnud <- dur.events(elevation = logger_PEGelev + .02, level = openedfile$ht_above_SL, units = "10 minutes", percentile = 0.5)
  
wet_filter <- openedfile %>%
  filter(sensors > 0 )
 
 wet_percentile = quantile(wet_filter$fixed_sensor, probs = c(.75)) 
 
 
eighty_per_depth = quantile(openedfile$fixed_sensor, probs = c(.8))
 ninetyfive_per_depth = quantile(openedfile$fixed_sensor, probs = c(.95))
 
 med_depth <-fld.depth(level = openedfile$ht_above_SL, elevation=logger_PEGelev + .02, percentile = 0.5)
flood_table_PEG <- flood_table_PEG %>%
  add_row(Serial_Number = as.integer(serialnum), Proportion_Time_Innudated = at_loggerelev_freq, `Median_Time_Innudated` = timeinnud, `Median_Depth_Innudated` = med_depth, `Logger_Elevation` = logger_PEGelev, `Type` = logger_PEGtype, `percentile_80_depth` =  eighty_per_depth, `percentile_95_depth` = ninetyfive_per_depth, `wet_75_percentile_depth` = wet_percentile)
 


}
```

```{r}
flood_table_PEG_Platonly <- flood_table_PEG %>%
  filter(Type == "Platform"| Type == "Edge" | Type == "Edge/Shrub") %>%
  filter(Serial_Number != "20389386")
flood_table_PEG_Platonly %>%
  ggplot(aes(Logger_Elevation, Proportion_Time_Innudated)) + geom_point()+ geom_smooth(method = lm) + ggtitle("Platform only")
flood_table_PEG_Platonly %>%
  ggplot(aes(Logger_Elevation, Median_Time_Innudated)) + geom_point()+ geom_smooth(method = lm) + ggtitle("Platform only")

summary(lm(Proportion_Time_Innudated ~ Logger_Elevation, flood_table_PEG_Platonly))

summary(lm( Median_Time_Innudated ~ Logger_Elevation, flood_table_PEG_Platonly))
```

```{r}
#flood_table_PEG_Platonly %>%
#  ggplot(aes(Logger_Elevation, ninetyfive_per_depth)) + geom_point()+ #geom_smooth(method = lm) + ggtitle("Platform only, 95th percentile depth")




summary(lm(Proportion_Time_Innudated ~ Logger_Elevation, flood_table_PEG_Platonly))


flood_table_PEG_Platonly %>%
  ggplot(aes(Logger_Elevation, Median_Time_Innudated)) + geom_point() + 
  geom_smooth(method = lm)



flood_table_PEG_PlatDitch<- flood_table_PEG %>%
  filter(Type == "Plat" | Type == "Ditch" | Type == "Edge") 
flood_table_PEG_PlatDitch %>%
  ggplot(aes(Logger_Elevation,  Proportion_Time_Innudated)) + geom_point() + 
  geom_smooth(method = lm) + ggtitle("Platform and Ditch")

flood_table_PEG_PlatDitchCreek <- flood_table_PEG %>%
  filter(Type == "Plat" | Type == "Ditch" | Type == "Creek"| Type == "Edge") 
flood_table_PEG_PlatDitchCreek %>%
  ggplot(aes(Logger_Elevation,  Proportion_Time_Innudated)) + geom_point() + 
  geom_smooth(method = lm) + ggtitle("Platform, Ditch and Creek")


#Only completed up to here so far


flood_table_PEG_PlatDitchCreek %>%
  ggplot(aes(Logger_Elevation, percentile_80_depth)) + geom_point() + 
  geom_smooth(method = lm) + ggtitle("Westport (no ponds) 80% depth")
flood_table_PEG_Platonly %>%
  ggplot(aes(Logger_Elevation, percentile_95_depth)) + geom_point() + 
  geom_smooth(method = lm) + ggtitle("Westport Platform 95% depth")
#The loggers were deployed for about 21 days, this graph represents the min depth for 50 hours of the 21 days. For 50 hours, the depth of the water was this depth or higher. 


flood_table_PEG_PlatDitchCreek %>%
  ggplot(aes(Logger_Elevation, wet_75_percentile_depth)) + geom_point() + 
  geom_smooth(method = lm) + ggtitle("Westport (no ponds) 75% depth (only wet data)")
```

Reading in the channel metric datasheet created in QGIS Cleaning the datatable into only the columns I want.

```{r}
wes_channel_metrics <- read_csv("/Users/evagerstle/Downloads/WES_logger_all_metrics_table.csv")

wes_channel_metrics <- wes_channel_metrics %>% 
  mutate(ID = field_1, Serial_Number = field_2, Elevation = field_5) %>%
  select(ID, Serial_Number, Elevation, H_OFD_Dis1, H_OFD_Out_, EucDisSt_1, EucDisOut_)



```

Joining channel datasheet with the floodtable sheet by the key of Serial Number

```{r}

wes_channel_metrics_flood_table_Platonly <- left_join(flood_table_WES_Platonly, wes_channel_metrics,by = join_by(Serial_Number))

wes_channel_metrics_flood_table_Platonly <- wes_channel_metrics_flood_table_Platonly %>% 
  mutate(OutDisDif = H_OFD_Out_- EucDisOut_) %>%
  mutate(ChaDisDif = H_OFD_Dis1 - EucDisSt_1)

```

Run initial simple Linear Regression

```{r}
Wes_prop_mod <- lm(Proportion_Time_Innudated ~ Logger_Elevation, data= wes_channel_metrics_flood_table_Platonly)
summary(Wes_prop_mod)


Wes_time_mod <- lm(Median_Time_Innudated ~ Logger_Elevation, data= wes_channel_metrics_flood_table_Platonly)
summary(Wes_Time_mod)


Wes_depth_mod <- lm(percentile_95_depth ~ Logger_Elevation, data= wes_channel_metrics_flood_table_Platonly)
summary(Wes_depth_mod)
```